---
title: "Final Project - STA 652"
author: "Richel Attafuah"
date: "2026-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coherent Hourly Traffic Flow Forecasting: Hierarchical Modeling and Reconciliation in R

## Reproducible workflow and leakage control

This project is implemented entirely in R using the `fpp3` ecosystem. The workflow is designed to be methodologically rigorous and reproducible.
A central design principle is the prevention of data leakage. The data are split into training and test periods before any imputation used for model fitting is performed. Missing values in the training set are imputed using summaries computed from the training data only. The test target values are never imputed. Instead, the evaluation sample is restricted to stations with fully observed test outcomes so that holdout accuracy is measured against genuine observed traffic counts.

To keep the analysis modular and reproducible, I define a small set of helper functions for hierarchy labeling, accuracy summarization, model fitting, and leakage-safe split preparation.

```{r include=FALSE}
############################################################
# 0. Setup and helper functions
############################################################

install.packages(c("fpp3", "lubridate", "patchwork"))

library(fpp3)
library(lubridate)
library(patchwork)

# ---------------------------
# Project paths
# ---------------------------
# Recommended repository structure:
# coherent-traffic-forecasting-r/
#   data/cleaned_traffic_data.csv
#   scripts/
#   outputs/
#   report/

data_path <- "data/cleaned_traffic_data.csv"

# ---------------------------
# Accuracy measures used throughout
# ---------------------------
accuracy_measures <- list(
  RMSE  = RMSE,
  MAE   = MAE,
  MASE  = MASE,
  RMSSE = RMSSE
)

# ---------------------------
# Helper: add hierarchy level labels
# ---------------------------
add_hierarchy_level <- function(x) {
  x |>
    mutate(
      hierarchy_level = case_when(
        is_aggregated(route) &
          is_aggregated(direction) &
          is_aggregated(lane_type) &
          is_aggregated(station) ~ "Total",

        !is_aggregated(route) &
          is_aggregated(direction) &
          is_aggregated(lane_type) &
          is_aggregated(station) ~ "Route",

        !is_aggregated(route) &
          !is_aggregated(direction) &
          is_aggregated(lane_type) &
          is_aggregated(station) ~ "Route-Direction",

        !is_aggregated(route) &
          !is_aggregated(direction) &
          !is_aggregated(lane_type) &
          is_aggregated(station) ~ "Route-Direction-LaneType",

        !is_aggregated(route) &
          !is_aggregated(direction) &
          !is_aggregated(lane_type) &
          !is_aggregated(station) ~ "Station",

        TRUE ~ "Other"
      )
    )
}

# ---------------------------
# Helper: summarise accuracy by hierarchy level
# ---------------------------
summarise_accuracy_by_level <- function(acc_tbl) {
  acc_tbl |>
    add_hierarchy_level() |>
    group_by(.model, hierarchy_level) |>
    summarise(
      RMSE  = mean(RMSE, na.rm = TRUE),
      MAE   = mean(MAE, na.rm = TRUE),
      MASE  = mean(MASE, na.rm = TRUE),
      RMSSE = mean(RMSSE, na.rm = TRUE),
      .groups = "drop"
    ) |>
    arrange(hierarchy_level, RMSSE)
}

# ---------------------------
# Helper: fit base models
# ---------------------------
fit_base_suite <- function(train_hier) {
  train_hier |>
    model(
      ets   = ETS(flow),
      arima = ARIMA(flow)
    )
}

# ---------------------------
# Helper: fit reconciliation suite
# ---------------------------
fit_reconciliation_suite <- function(base_models) {
  base_models |>
    reconcile(
      ets_bu            = bottom_up(ets),
      ets_td_forecast   = top_down(ets, method = "forecast_proportions"),
      ets_td_historical = top_down(ets, method = "average_proportions"),
      ets_mint_shrink   = min_trace(ets, method = "mint_shrink"),
      ets_mint_sample   = min_trace(ets, method = "mint_cov"),

      arima_bu            = bottom_up(arima),
      arima_td_forecast   = top_down(arima, method = "forecast_proportions"),
      arima_td_historical = top_down(arima, method = "average_proportions"),
      arima_mint_shrink   = min_trace(arima, method = "mint_shrink"),
      arima_mint_sample   = min_trace(arima, method = "mint_cov")
    )
}

# ---------------------------
# Helper: leakage-safe split preparation
# ---------------------------
# This function:
# 1. splits the balanced bottom-level panel into train/test,
# 2. computes imputation references using TRAIN ONLY,
# 3. imputes TRAIN ONLY,
# 4. leaves TEST actuals untouched,
# 5. returns hierarchical train/test tsibbles.
prepare_hier_split <- function(panel_data, cutoff_time) {

  train_raw <- panel_data |>
    filter(timestamp <= cutoff_time) |>
    mutate(
      day_of_week  = wday(timestamp, week_start = 1),
      hour_of_day  = hour(timestamp),
      hour_of_week = (day_of_week - 1L) * 24L + hour_of_day
    )

  test_raw <- panel_data |>
    filter(timestamp > cutoff_time)

  observed_train <- train_raw |>
    filter(!is.na(total_flow))

  if (nrow(observed_train) == 0) {
    stop("No observed training values are available after filtering.")
  }

  # Training-only imputation references
  ref_station_hw <- observed_train |>
    group_by(station, hour_of_week) |>
    summarise(ref_station_hw = median(total_flow), .groups = "drop")

  ref_station <- observed_train |>
    group_by(station) |>
    summarise(ref_station = median(total_flow), .groups = "drop")

  ref_cell_hw <- observed_train |>
    group_by(route, direction, lane_type, hour_of_week) |>
    summarise(ref_cell_hw = median(total_flow), .groups = "drop")

  ref_cell <- observed_train |>
    group_by(route, direction, lane_type) |>
    summarise(ref_cell = median(total_flow), .groups = "drop")

  ref_global_hw <- observed_train |>
    group_by(hour_of_week) |>
    summarise(ref_global_hw = median(total_flow), .groups = "drop")

  ref_global <- median(observed_train$total_flow, na.rm = TRUE)

  # Impute TRAIN ONLY using training-derived references
  train_imputed <- train_raw |>
    left_join(ref_station_hw, by = c("station", "hour_of_week")) |>
    left_join(ref_station, by = "station") |>
    left_join(ref_cell_hw, by = c("route", "direction", "lane_type", "hour_of_week")) |>
    left_join(ref_cell, by = c("route", "direction", "lane_type")) |>
    left_join(ref_global_hw, by = "hour_of_week") |>
    mutate(
      flow = case_when(
        !is.na(total_flow) ~ total_flow,
        is.na(total_flow) & !is.na(ref_station_hw) ~ ref_station_hw,
        is.na(total_flow) & !is.na(ref_station) ~ ref_station,
        is.na(total_flow) & !is.na(ref_cell_hw) ~ ref_cell_hw,
        is.na(total_flow) & !is.na(ref_cell) ~ ref_cell,
        is.na(total_flow) & !is.na(ref_global_hw) ~ ref_global_hw,
        TRUE ~ ref_global
      )
    ) |>
    select(timestamp, station, route, direction, lane_type, flow)

  if (any(is.na(train_imputed$flow))) {
    stop("Training imputation produced NA values. Check the reference cascade.")
  }

  # Never impute the test target
  if (any(is.na(test_raw$total_flow))) {
    stop("Test set contains missing actuals. Revise the station screen so evaluation remains fully observed.")
  }

  test_actuals <- test_raw |>
    transmute(
      timestamp,
      station,
      route,
      direction,
      lane_type,
      flow = total_flow
    )

  train_bottom <- train_imputed |>
    as_tsibble(key = c(route, direction, lane_type, station), index = timestamp)

  test_bottom <- test_actuals |>
    as_tsibble(key = c(route, direction, lane_type, station), index = timestamp)

  train_hier <- train_bottom |>
    aggregate_key(
      route / direction / lane_type / station,
      flow = sum(flow)
    )

  test_hier <- test_bottom |>
    aggregate_key(
      route / direction / lane_type / station,
      flow = sum(flow)
    )

  list(
    train_bottom = train_bottom,
    test_bottom  = test_bottom,
    train_hier   = train_hier,
    test_hier    = test_hier
  )
}
```


# 1. Problem formulation and hierarchy definition

## 1.1 Forecasting objective

This project studies hierarchical time series forecasting and forecast reconciliation using hourly traffic flow data. The forecasting problem is inherently multilevel: station-level flows aggregate to route-direction-lane-type totals, which aggregate further to route-direction totals, then to route totals, and ultimately to the system total. If forecasts are generated independently at each level, they will generally be incoherent, meaning that lower-level forecasts will not sum exactly to higher-level forecasts.

The aim of the project is therefore to compare alternative reconciliation strategies that transform independently generated base forecasts into a coherent forecasting system.

## 1.2 Why this dataset is suitable

The dataset contains regularly spaced hourly observations and a large number of bottom-level traffic stations. The core structural variables are `Route`, `Direction of Travel`, `Lane Type`, and `Station`. These fields support a clear additive hierarchy. The `District` field is not used in the hierarchy because it does not provide meaningful variation in the uploaded extract.

## 1.3 Main hierarchy

The main hierarchy used in this project is:

\[
\text{Total}
\rightarrow
\text{Route}
\rightarrow
\text{Route-Direction}
\rightarrow
\text{Route-Direction-Lane Type}
\rightarrow
\text{Station}.
\]

This design satisfies the assignment requirement of having at least three hierarchical levels and provides a natural operational interpretation for traffic forecasting.


```{r echo=FALSE}
############################################################
# Problem formulation and data audit
############################################################

traffic_raw <- readr::read_csv("C:\\Users\\attafuro\\Desktop\\EV Charging Analysis\\STA 652 Project\\cleaned_traffic_data.csv", show_col_types = FALSE) |>
  transmute(
    timestamp = parse_date_time(
      Timestamp,
      orders = c("mdy HM", "mdy HMS"),
      tz = "UTC"
    ),
    station        = as.character(Station),
    district       = as.character(District),
    route          = as.character(Route),
    direction      = as.character(`Direction of Travel`),
    lane_type      = as.character(`Lane Type`),
    station_length = as.numeric(`Station Length`),
    samples        = as.numeric(Samples),
    pct_observed   = as.numeric(`% Observed`),
    total_flow     = as.numeric(`Total Flow`)
  ) |>
  arrange(station, timestamp)

# High-level data profile
data_profile <- traffic_raw |>
  summarise(
    n_rows        = n(),
    n_timestamps  = n_distinct(timestamp),
    n_stations    = n_distinct(station),
    n_routes      = n_distinct(route),
    n_directions  = n_distinct(direction),
    n_lane_types  = n_distinct(lane_type),
    n_districts   = n_distinct(district),
    min_time      = min(timestamp, na.rm = TRUE),
    max_time      = max(timestamp, na.rm = TRUE)
  )

print(data_profile)

# Check whether station metadata are structurally consistent
station_mapping_check <- traffic_raw |>
  group_by(station) |>
  summarise(
    n_routes     = n_distinct(route),
    n_directions = n_distinct(direction),
    n_lane_types = n_distinct(lane_type),
    .groups = "drop"
  )

stopifnot(
  max(station_mapping_check$n_routes) == 1,
  max(station_mapping_check$n_directions) == 1,
  max(station_mapping_check$n_lane_types) == 1
)

# Check duplicate station-time observations
duplicate_check <- traffic_raw |>
  count(station, timestamp, name = "n") |>
  filter(n > 1)

stopifnot(nrow(duplicate_check) == 0)

# Keep one clean station metadata table
station_meta <- traffic_raw |>
  distinct(station, route, direction, lane_type)
```


# 2. Leakage-safe sample construction

## 2.1 Why the split comes before imputation

To avoid leakage, the forecasting origin is defined before any training imputation is carried out. This means that any missing values used for model fitting are imputed using only information available within the training period. The test target values are never imputed.

This choice is important because using the full sample to estimate imputation templates would allow future information to influence model training. That would bias forecast evaluation.

## 2.2 Balanced panel construction

The raw file may contain implicit gaps, where a station-hour combination is absent rather than explicitly recorded as missing. To make the forecasting panel well defined, I first construct the full station-by-hour grid and then join the raw observations back onto that grid. This makes all missingness explicit.

## 2.3 Station screening

The final evaluation sample is selected using the following rules:

1. the station must be present for the full hourly training and test index,
2. the station must have no missing test target values,
3. the training missingness rate must be low.

The second rule is used only to guarantee a valid holdout evaluation sample. It does not provide the model with future target values, and the test outcomes themselves remain untouched.

```{r echo=FALSE}
############################################################
#  Leakage-safe sample construction and train/test split
############################################################

# Build a balanced station-by-hour panel so that implicit gaps become explicit
all_timestamps <- traffic_raw |>
  distinct(timestamp) |>
  arrange(timestamp)

traffic_panel <- station_meta |>
  tidyr::crossing(timestamp = all_timestamps$timestamp) |>
  left_join(
    traffic_raw |>
      select(timestamp, station, route, direction, lane_type, pct_observed, total_flow),
    by = c("timestamp", "station", "route", "direction", "lane_type")
  ) |>
  arrange(station, timestamp)

# Main holdout horizon: last 14 days = 336 hours
h_test <- 24 * 14
cutoff_time <- max(traffic_panel$timestamp) - hours(h_test)

n_train_times <- traffic_panel |>
  filter(timestamp <= cutoff_time) |>
  distinct(timestamp) |>
  nrow()

n_test_times <- traffic_panel |>
  filter(timestamp > cutoff_time) |>
  distinct(timestamp) |>
  nrow()

# Screen stations for a leakage-safe analysis sample
station_screen <- traffic_panel |>
  mutate(segment = if_else(timestamp <= cutoff_time, "train", "test")) |>
  group_by(station, route, direction, lane_type, segment) |>
  summarise(
    n_rows             = n(),
    missing_flow_n     = sum(is.na(total_flow)),
    missing_flow_share = mean(is.na(total_flow)),
    mean_pct_observed  = mean(pct_observed, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = segment,
    values_from = c(n_rows, missing_flow_n, missing_flow_share, mean_pct_observed),
    names_sep = "_"
  )

study_stations <- station_screen |>
  filter(
    n_rows_train == n_train_times,
    n_rows_test == n_test_times,
    missing_flow_n_test == 0,
    missing_flow_share_train <= 0.05
  ) |>
  select(station, route, direction, lane_type)

# Assignment requires at least 100 bottom-level series
stopifnot(nrow(study_stations) >= 100)

traffic_study <- traffic_panel |>
  semi_join(study_stations, by = c("station", "route", "direction", "lane_type"))

analysis_sample_summary <- traffic_study |>
  summarise(
    bottom_series = n_distinct(station),
    timestamps    = n_distinct(timestamp),
    routes        = n_distinct(route),
    directions    = n_distinct(direction),
    lane_types    = n_distinct(lane_type)
  )

print(analysis_sample_summary)
```

# 3. Balanced bottom-level panel and hierarchy construction

## 3.1 Training-only imputation

Any missing training values are imputed using a cascade of training-only reference summaries:

1. station × hour-of-week median,
2. station median,
3. route-direction-lane-type × hour-of-week median,
4. route-direction-lane-type median,
5. global hour-of-week median,
6. global median.

This is intentionally conservative. The goal is not to create synthetic signal, but to repair a small number of training gaps while preserving recurring intraday and intraweek structure.

## 3.2 Why the test set is not imputed

The test set is reserved for genuine out-of-sample evaluation. Imputing test target values would contaminate the accuracy assessment. Therefore, the evaluation sample is restricted to stations with fully observed test outcomes.

## 3.3 Hierarchical representation

Let \(\mathbf{b}_t\) denote the vector of bottom-level station flows and let \(\mathbf{y}_t\) denote the full vector of all series in the hierarchy. Then

\[
\mathbf{y}_t = \mathbf{S}\mathbf{b}_t,
\]

where \(\mathbf{S}\) is the summing matrix. Each row of \(\mathbf{S}\) corresponds to either an aggregate series or a bottom-level series, and the matrix encodes the additive structure of the hierarchy.


```{r echo=FALSE}
############################################################
# Training-only imputation and hierarchy construction
############################################################

main_split <- prepare_hier_split(
  panel_data  = traffic_study,
  cutoff_time = cutoff_time
)

train_bottom <- main_split$train_bottom
test_bottom  <- main_split$test_bottom
train_hier   <- main_split$train_hier
test_hier    <- main_split$test_hier

# Forecast horizon implied by the holdout
h_main <- test_hier |>
  distinct(timestamp) |>
  nrow()

# Quick structural summary
hierarchy_sample_summary <- tibble(
  bottom_series = n_distinct(train_bottom$station),
  train_hours   = n_distinct(train_bottom$timestamp),
  test_hours    = n_distinct(test_bottom$timestamp),
  routes        = n_distinct(train_bottom$route),
  directions    = n_distinct(train_bottom$direction),
  lane_types    = n_distinct(train_bottom$lane_type)
)

print(hierarchy_sample_summary)
```


# Exploratory analysis

All exploratory analysis shown below is based on the training period only. This keeps the descriptive work consistent with the leakage-control strategy and avoids using the holdout data to inform modeling decisions.

At the aggregate level, I examine the top-level series to assess overall trend, variability, and seasonality. I also inspect average hourly and day-of-week profiles to determine whether the traffic system exhibits strong intraday and intraweek structure. These diagnostics help justify the use of standard univariate forecasting models such as ETS and ARIMA.


```{r echo=FALSE}
############################################################
#  Exploratory analysis using training data only
############################################################

# Top-level total series
train_total <- train_hier |>
  filter(
    is_aggregated(route),
    is_aggregated(direction),
    is_aggregated(lane_type),
    is_aggregated(station)
  )

test_total <- test_hier |>
  filter(
    is_aggregated(route),
    is_aggregated(direction),
    is_aggregated(lane_type),
    is_aggregated(station)
  )

p_total <- bind_rows(train_total, test_total) |>
  autoplot(flow) +
  geom_vline(xintercept = cutoff_time, linetype = "dashed") +
  labs(
    title = "Top-level hourly traffic flow",
    subtitle = "Dashed line marks the forecast origin",
    x = "Time",
    y = "Flow"
  )

# Mean hourly profile from TRAIN only
hour_profile <- train_bottom |>
  as_tibble() |>
  mutate(hour = hour(timestamp)) |>
  group_by(hour) |>
  summarise(mean_flow = mean(flow), .groups = "drop")

p_hour <- hour_profile |>
  ggplot(aes(x = hour, y = mean_flow)) +
  geom_line() +
  labs(
    title = "Mean hourly traffic profile (training data)",
    x = "Hour of day",
    y = "Mean flow"
  )

# Mean day-of-week profile from TRAIN only
dow_profile <- train_bottom |>
  as_tibble() |>
  mutate(day = wday(timestamp, label = TRUE, week_start = 1)) |>
  group_by(day) |>
  summarise(mean_flow = mean(flow), .groups = "drop")

p_dow <- dow_profile |>
  ggplot(aes(x = day, y = mean_flow, group = 1)) +
  geom_line() +
  labs(
    title = "Mean day-of-week traffic profile (training data)",
    x = "Day of week",
    y = "Mean flow"
  )

p_total / (p_hour | p_dow)
```

