---
title: "Final Project - STA 652"
author: "Richel Attafuah"
date: "2026-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coherent Hourly Traffic Flow Forecasting: Hierarchical Modeling and Reconciliation in R

## Reproducible workflow and leakage control

This project is implemented entirely in R using the `fpp3` ecosystem. The workflow is designed to be methodologically rigorous and reproducible.
A central design principle is the prevention of data leakage. The data are split into training and test periods before any imputation used for model fitting is performed. Missing values in the training set are imputed using summaries computed from the training data only. The test target values are never imputed. Instead, the evaluation sample is restricted to stations with fully observed test outcomes so that holdout accuracy is measured against genuine observed traffic counts.

To keep the analysis modular and reproducible, I define a small set of helper functions for hierarchy labeling, accuracy summarization, model fitting, and leakage-safe split preparation.

```{r include=FALSE}
############################################################
# 0. Setup and helper functions
############################################################

install.packages(c("fpp3", "lubridate", "patchwork"))

library(fpp3)
library(lubridate)
library(patchwork)

# ---------------------------
# Project paths
# ---------------------------
# Recommended repository structure:
# coherent-traffic-forecasting-r/
#   data/cleaned_traffic_data.csv
#   scripts/
#   outputs/
#   report/

data_path <- "data/cleaned_traffic_data.csv"

# ---------------------------
# Accuracy measures used throughout
# ---------------------------
accuracy_measures <- list(
  RMSE  = RMSE,
  MAE   = MAE,
  MASE  = MASE,
  RMSSE = RMSSE
)

# ---------------------------
# Helper: add hierarchy level labels
# ---------------------------
add_hierarchy_level <- function(x) {
  x |>
    mutate(
      hierarchy_level = case_when(
        is_aggregated(route) &
          is_aggregated(direction) &
          is_aggregated(lane_type) &
          is_aggregated(station) ~ "Total",

        !is_aggregated(route) &
          is_aggregated(direction) &
          is_aggregated(lane_type) &
          is_aggregated(station) ~ "Route",

        !is_aggregated(route) &
          !is_aggregated(direction) &
          is_aggregated(lane_type) &
          is_aggregated(station) ~ "Route-Direction",

        !is_aggregated(route) &
          !is_aggregated(direction) &
          !is_aggregated(lane_type) &
          is_aggregated(station) ~ "Route-Direction-LaneType",

        !is_aggregated(route) &
          !is_aggregated(direction) &
          !is_aggregated(lane_type) &
          !is_aggregated(station) ~ "Station",

        TRUE ~ "Other"
      )
    )
}

# ---------------------------
# Helper: summarise accuracy by hierarchy level
# ---------------------------
summarise_accuracy_by_level <- function(acc_tbl) {
  acc_tbl |>
    add_hierarchy_level() |>
    group_by(.model, hierarchy_level) |>
    summarise(
      RMSE  = mean(RMSE, na.rm = TRUE),
      MAE   = mean(MAE, na.rm = TRUE),
      MASE  = mean(MASE, na.rm = TRUE),
      RMSSE = mean(RMSSE, na.rm = TRUE),
      .groups = "drop"
    ) |>
    arrange(hierarchy_level, RMSSE)
}

# ---------------------------
# Helper: fit base models
# ---------------------------
fit_base_suite <- function(train_hier) {
  train_hier |>
    model(
      ets   = ETS(flow),
      arima = ARIMA(flow)
    )
}

# ---------------------------
# Helper: fit reconciliation suite
# ---------------------------
fit_reconciliation_suite <- function(base_models) {
  base_models |>
    reconcile(
      ets_bu            = bottom_up(ets),
      ets_td_forecast   = top_down(ets, method = "forecast_proportions"),
      ets_td_historical = top_down(ets, method = "average_proportions"),
      ets_mint_shrink   = min_trace(ets, method = "mint_shrink"),
      ets_mint_sample   = min_trace(ets, method = "mint_cov"),

      arima_bu            = bottom_up(arima),
      arima_td_forecast   = top_down(arima, method = "forecast_proportions"),
      arima_td_historical = top_down(arima, method = "average_proportions"),
      arima_mint_shrink   = min_trace(arima, method = "mint_shrink"),
      arima_mint_sample   = min_trace(arima, method = "mint_cov")
    )
}

# ---------------------------
# Helper: leakage-safe split preparation
# ---------------------------
# This function:
# 1. splits the balanced bottom-level panel into train/test,
# 2. computes imputation references using TRAIN ONLY,
# 3. imputes TRAIN ONLY,
# 4. leaves TEST actuals untouched,
# 5. returns hierarchical train/test tsibbles.
prepare_hier_split <- function(panel_data, cutoff_time) {

  train_raw <- panel_data |>
    filter(timestamp <= cutoff_time) |>
    mutate(
      day_of_week  = wday(timestamp, week_start = 1),
      hour_of_day  = hour(timestamp),
      hour_of_week = (day_of_week - 1L) * 24L + hour_of_day
    )

  test_raw <- panel_data |>
    filter(timestamp > cutoff_time)

  observed_train <- train_raw |>
    filter(!is.na(total_flow))

  if (nrow(observed_train) == 0) {
    stop("No observed training values are available after filtering.")
  }

  # Training-only imputation references
  ref_station_hw <- observed_train |>
    group_by(station, hour_of_week) |>
    summarise(ref_station_hw = median(total_flow), .groups = "drop")

  ref_station <- observed_train |>
    group_by(station) |>
    summarise(ref_station = median(total_flow), .groups = "drop")

  ref_cell_hw <- observed_train |>
    group_by(route, direction, lane_type, hour_of_week) |>
    summarise(ref_cell_hw = median(total_flow), .groups = "drop")

  ref_cell <- observed_train |>
    group_by(route, direction, lane_type) |>
    summarise(ref_cell = median(total_flow), .groups = "drop")

  ref_global_hw <- observed_train |>
    group_by(hour_of_week) |>
    summarise(ref_global_hw = median(total_flow), .groups = "drop")

  ref_global <- median(observed_train$total_flow, na.rm = TRUE)

  # Impute TRAIN ONLY using training-derived references
  train_imputed <- train_raw |>
    left_join(ref_station_hw, by = c("station", "hour_of_week")) |>
    left_join(ref_station, by = "station") |>
    left_join(ref_cell_hw, by = c("route", "direction", "lane_type", "hour_of_week")) |>
    left_join(ref_cell, by = c("route", "direction", "lane_type")) |>
    left_join(ref_global_hw, by = "hour_of_week") |>
    mutate(
      flow = case_when(
        !is.na(total_flow) ~ total_flow,
        is.na(total_flow) & !is.na(ref_station_hw) ~ ref_station_hw,
        is.na(total_flow) & !is.na(ref_station) ~ ref_station,
        is.na(total_flow) & !is.na(ref_cell_hw) ~ ref_cell_hw,
        is.na(total_flow) & !is.na(ref_cell) ~ ref_cell,
        is.na(total_flow) & !is.na(ref_global_hw) ~ ref_global_hw,
        TRUE ~ ref_global
      )
    ) |>
    select(timestamp, station, route, direction, lane_type, flow)

  if (any(is.na(train_imputed$flow))) {
    stop("Training imputation produced NA values. Check the reference cascade.")
  }

  # Never impute the test target
  if (any(is.na(test_raw$total_flow))) {
    stop("Test set contains missing actuals. Revise the station screen so evaluation remains fully observed.")
  }

  test_actuals <- test_raw |>
    transmute(
      timestamp,
      station,
      route,
      direction,
      lane_type,
      flow = total_flow
    )

  train_bottom <- train_imputed |>
    as_tsibble(key = c(route, direction, lane_type, station), index = timestamp)

  test_bottom <- test_actuals |>
    as_tsibble(key = c(route, direction, lane_type, station), index = timestamp)

  train_hier <- train_bottom |>
    aggregate_key(
      route / direction / lane_type / station,
      flow = sum(flow)
    )

  test_hier <- test_bottom |>
    aggregate_key(
      route / direction / lane_type / station,
      flow = sum(flow)
    )

  list(
    train_bottom = train_bottom,
    test_bottom  = test_bottom,
    train_hier   = train_hier,
    test_hier    = test_hier
  )
}
```


# 1. Problem formulation and hierarchy definition

## 1.1 Forecasting objective

This project studies hierarchical time series forecasting and forecast reconciliation using hourly traffic flow data. The forecasting problem is inherently multilevel: station-level flows aggregate to route-direction-lane-type totals, which aggregate further to route-direction totals, then to route totals, and ultimately to the system total. If forecasts are generated independently at each level, they will generally be incoherent, meaning that lower-level forecasts will not sum exactly to higher-level forecasts.

The aim of the project is therefore to compare alternative reconciliation strategies that transform independently generated base forecasts into a coherent forecasting system.

## 1.2 Why this dataset is suitable

The dataset contains regularly spaced hourly observations and a large number of bottom-level traffic stations. The core structural variables are `Route`, `Direction of Travel`, `Lane Type`, and `Station`. These fields support a clear additive hierarchy. The `District` field is not used in the hierarchy because it does not provide meaningful variation in the uploaded extract.

## 1.3 Main hierarchy

The main hierarchy used in this project is:

\[
\text{Total}
\rightarrow
\text{Route}
\rightarrow
\text{Route-Direction}
\rightarrow
\text{Route-Direction-Lane Type}
\rightarrow
\text{Station}.
\]

This design satisfies the assignment requirement of having at least three hierarchical levels and provides a natural operational interpretation for traffic forecasting.


```{r echo=FALSE}
############################################################
# Problem formulation and data audit
############################################################

traffic_raw <- readr::read_csv("C:\\Users\\attafuro\\Desktop\\EV Charging Analysis\\STA 652 Project\\cleaned_traffic_data.csv", show_col_types = FALSE) |>
  transmute(
    timestamp = parse_date_time(
      Timestamp,
      orders = c("mdy HM", "mdy HMS"),
      tz = "UTC"
    ),
    station        = as.character(Station),
    district       = as.character(District),
    route          = as.character(Route),
    direction      = as.character(`Direction of Travel`),
    lane_type      = as.character(`Lane Type`),
    station_length = as.numeric(`Station Length`),
    samples        = as.numeric(Samples),
    pct_observed   = as.numeric(`% Observed`),
    total_flow     = as.numeric(`Total Flow`)
  ) |>
  arrange(station, timestamp)

# High-level data profile
data_profile <- traffic_raw |>
  summarise(
    n_rows        = n(),
    n_timestamps  = n_distinct(timestamp),
    n_stations    = n_distinct(station),
    n_routes      = n_distinct(route),
    n_directions  = n_distinct(direction),
    n_lane_types  = n_distinct(lane_type),
    n_districts   = n_distinct(district),
    min_time      = min(timestamp, na.rm = TRUE),
    max_time      = max(timestamp, na.rm = TRUE)
  )

print(data_profile)

# Check whether station metadata are structurally consistent
station_mapping_check <- traffic_raw |>
  group_by(station) |>
  summarise(
    n_routes     = n_distinct(route),
    n_directions = n_distinct(direction),
    n_lane_types = n_distinct(lane_type),
    .groups = "drop"
  )

stopifnot(
  max(station_mapping_check$n_routes) == 1,
  max(station_mapping_check$n_directions) == 1,
  max(station_mapping_check$n_lane_types) == 1
)

# Check duplicate station-time observations
duplicate_check <- traffic_raw |>
  count(station, timestamp, name = "n") |>
  filter(n > 1)

stopifnot(nrow(duplicate_check) == 0)

# Keep one clean station metadata table
station_meta <- traffic_raw |>
  distinct(station, route, direction, lane_type)
```


# 2. Leakage-safe sample construction

## 2.1 Why the split comes before imputation

To avoid leakage, the forecasting origin is defined before any training imputation is carried out. This means that any missing values used for model fitting are imputed using only information available within the training period. The test target values are never imputed.

This choice is important because using the full sample to estimate imputation templates would allow future information to influence model training. That would bias forecast evaluation.

## 2.2 Balanced panel construction

The raw file may contain implicit gaps, where a station-hour combination is absent rather than explicitly recorded as missing. To make the forecasting panel well defined, I first construct the full station-by-hour grid and then join the raw observations back onto that grid. This makes all missingness explicit.

## 2.3 Station screening

The final evaluation sample is selected using the following rules:

1. the station must be present for the full hourly training and test index,
2. the station must have no missing test target values,
3. the training missingness rate must be low.

The second rule is used only to guarantee a valid holdout evaluation sample. It does not provide the model with future target values, and the test outcomes themselves remain untouched.

```{r echo=FALSE}
############################################################
#  Leakage-safe sample construction and train/test split
############################################################

# Build a balanced station-by-hour panel so that implicit gaps become explicit
all_timestamps <- traffic_raw |>
  distinct(timestamp) |>
  arrange(timestamp)

traffic_panel <- station_meta |>
  tidyr::crossing(timestamp = all_timestamps$timestamp) |>
  left_join(
    traffic_raw |>
      select(timestamp, station, route, direction, lane_type, pct_observed, total_flow),
    by = c("timestamp", "station", "route", "direction", "lane_type")
  ) |>
  arrange(station, timestamp)

# Main holdout horizon: last 14 days = 336 hours
h_test <- 24 * 14
cutoff_time <- max(traffic_panel$timestamp) - hours(h_test)

n_train_times <- traffic_panel |>
  filter(timestamp <= cutoff_time) |>
  distinct(timestamp) |>
  nrow()

n_test_times <- traffic_panel |>
  filter(timestamp > cutoff_time) |>
  distinct(timestamp) |>
  nrow()

# Screen stations for a leakage-safe analysis sample
station_screen <- traffic_panel |>
  mutate(segment = if_else(timestamp <= cutoff_time, "train", "test")) |>
  group_by(station, route, direction, lane_type, segment) |>
  summarise(
    n_rows             = n(),
    missing_flow_n     = sum(is.na(total_flow)),
    missing_flow_share = mean(is.na(total_flow)),
    mean_pct_observed  = mean(pct_observed, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = segment,
    values_from = c(n_rows, missing_flow_n, missing_flow_share, mean_pct_observed),
    names_sep = "_"
  )

study_stations <- station_screen |>
  filter(
    n_rows_train == n_train_times,
    n_rows_test == n_test_times,
    missing_flow_n_test == 0,
    missing_flow_share_train <= 0.05
  ) |>
  select(station, route, direction, lane_type)

# Assignment requires at least 100 bottom-level series
stopifnot(nrow(study_stations) >= 100)

traffic_study <- traffic_panel |>
  semi_join(study_stations, by = c("station", "route", "direction", "lane_type"))

analysis_sample_summary <- traffic_study |>
  summarise(
    bottom_series = n_distinct(station),
    timestamps    = n_distinct(timestamp),
    routes        = n_distinct(route),
    directions    = n_distinct(direction),
    lane_types    = n_distinct(lane_type)
  )

print(analysis_sample_summary)
```

# 3. Balanced bottom-level panel and hierarchy construction

## 3.1 Training-only imputation

Any missing training values are imputed using a cascade of training-only reference summaries:

1. station × hour-of-week median,
2. station median,
3. route-direction-lane-type × hour-of-week median,
4. route-direction-lane-type median,
5. global hour-of-week median,
6. global median.

This is intentionally conservative. The goal is not to create synthetic signal, but to repair a small number of training gaps while preserving recurring intraday and intraweek structure.

## 3.2 Why the test set is not imputed

The test set is reserved for genuine out-of-sample evaluation. Imputing test target values would contaminate the accuracy assessment. Therefore, the evaluation sample is restricted to stations with fully observed test outcomes.

## 3.3 Hierarchical representation

Let \(\mathbf{b}_t\) denote the vector of bottom-level station flows and let \(\mathbf{y}_t\) denote the full vector of all series in the hierarchy. Then

\[
\mathbf{y}_t = \mathbf{S}\mathbf{b}_t,
\]

where \(\mathbf{S}\) is the summing matrix. Each row of \(\mathbf{S}\) corresponds to either an aggregate series or a bottom-level series, and the matrix encodes the additive structure of the hierarchy.


```{r echo=FALSE}
############################################################
# Training-only imputation and hierarchy construction
############################################################

main_split <- prepare_hier_split(
  panel_data  = traffic_study,
  cutoff_time = cutoff_time
)

train_bottom <- main_split$train_bottom
test_bottom  <- main_split$test_bottom
train_hier   <- main_split$train_hier
test_hier    <- main_split$test_hier

# Forecast horizon implied by the holdout
h_main <- test_hier |>
  distinct(timestamp) |>
  nrow()

# Quick structural summary
hierarchy_sample_summary <- tibble(
  bottom_series = n_distinct(train_bottom$station),
  train_hours   = n_distinct(train_bottom$timestamp),
  test_hours    = n_distinct(test_bottom$timestamp),
  routes        = n_distinct(train_bottom$route),
  directions    = n_distinct(train_bottom$direction),
  lane_types    = n_distinct(train_bottom$lane_type)
)

print(hierarchy_sample_summary)
```


# Exploratory analysis

All exploratory analysis shown below is based on the training period only. This keeps the descriptive work consistent with the leakage-control strategy and avoids using the holdout data to inform modeling decisions.

At the aggregate level, I examine the top-level series to assess overall trend, variability, and seasonality. I also inspect average hourly and day-of-week profiles to determine whether the traffic system exhibits strong intraday and intraweek structure. These diagnostics help justify the use of standard univariate forecasting models such as ETS and ARIMA.


```{r echo=FALSE}
############################################################
#  Exploratory analysis using training data only
############################################################

# Top-level total series
train_total <- train_hier |>
  filter(
    is_aggregated(route),
    is_aggregated(direction),
    is_aggregated(lane_type),
    is_aggregated(station)
  )

test_total <- test_hier |>
  filter(
    is_aggregated(route),
    is_aggregated(direction),
    is_aggregated(lane_type),
    is_aggregated(station)
  )

p_total <- bind_rows(train_total, test_total) |>
  autoplot(flow) +
  geom_vline(xintercept = cutoff_time, linetype = "dashed") +
  labs(
    title = "Top-level hourly traffic flow",
    subtitle = "Dashed line marks the forecast origin",
    x = "Time",
    y = "Flow"
  )

# Mean hourly profile from TRAIN only
hour_profile <- train_bottom |>
  as_tibble() |>
  mutate(hour = hour(timestamp)) |>
  group_by(hour) |>
  summarise(mean_flow = mean(flow), .groups = "drop")

p_hour <- hour_profile |>
  ggplot(aes(x = hour, y = mean_flow)) +
  geom_line() +
  labs(
    title = "Mean hourly traffic profile (training data)",
    x = "Hour of day",
    y = "Mean flow"
  )

# Mean day-of-week profile from TRAIN only
dow_profile <- train_bottom |>
  as_tibble() |>
  mutate(day = wday(timestamp, label = TRUE, week_start = 1)) |>
  group_by(day) |>
  summarise(mean_flow = mean(flow), .groups = "drop")

p_dow <- dow_profile |>
  ggplot(aes(x = day, y = mean_flow, group = 1)) +
  geom_line() +
  labs(
    title = "Mean day-of-week traffic profile (training data)",
    x = "Day of week",
    y = "Mean flow"
  )

p_total / (p_hour | p_dow)
```


# 5. Base model specification

## 5.1 Purpose of the base models

The first modeling stage produces unconstrained base forecasts independently for every series in the hierarchy. These forecasts are not expected to be coherent. Reconciliation is applied afterward.

## 5.2 Chosen model classes

I fit two univariate model classes to each series:

1. ETS (Exponential Smoothing State Space Model),
2. ARIMA (Autoregressive Integrated Moving Average).


## 5.3 Why these models are appropriate

Traffic flow data typically exhibit serial dependence, recurrent patterns, and changing local level. ETS is useful because it captures evolving level, trend, and seasonal structure in a state space framework. ARIMA is useful because it models autocorrelation and differencing-based dynamics explicitly. Together, these two model classes provide a credible and interpretable base-model comparison.

```{r echo=FALSE}
############################################################
#  Base model specification
############################################################

base_models <- fit_base_suite(train_hier)

# Compact model diagnostics
base_model_glance <- glance(base_models)
print(base_model_glance)

# Base forecasts over the holdout horizon
fc_base <- base_models |>
  forecast(h = h_main)
```



# 6. Reconciliation theory and implementation

## 6.1 General reconciliation framework

Let \(\hat{\mathbf y}_h\) denote the vector of base forecasts at horizon \(h\). These forecasts are generally incoherent. Reconciled forecasts are constructed as

\[
\tilde{\mathbf y}_h = \mathbf{S}\mathbf{G}\hat{\mathbf y}_h,
\]

where \(\mathbf{S}\) is the summing matrix and \(\mathbf{G}\) maps the base forecasts into a coherent representation.

## 6.2 Reconciliation methods considered

I compare the following reconciliation strategies:

1. **Bottom-up**  
   Bottom-level forecasts are generated first and then aggregated upward.

2. **Top-down using forecast proportions**  
   The top-level forecast is allocated downward using proportions implied by the lower-level base forecasts.

3. **Top-down using historical proportions**  
   The top-level forecast is allocated downward using historical average shares.

4. **MinT with shrinkage covariance**  
   The reconciliation weights are chosen to minimize forecast error variance using a shrinkage estimate of the forecast error covariance matrix.

5. **MinT with sample covariance**  
   The reconciliation weights are based on the sample covariance matrix of residuals.


```{r echo=FALSE}
############################################################
# Reconciliation theory and implementation
############################################################

reconciled_models <- fit_reconciliation_suite(base_models)

# Reconciled forecasts over the holdout horizon
fc_reconciled <- reconciled_models |>
  forecast(h = h_main)

```


# 7. Forecast evaluation

## 7.1 Evaluation design

The final 14 days of the sample are used as a holdout period. Forecast accuracy is evaluated using both scale-dependent and scale-free measures:

- RMSE,
- MAE,
- MASE,
- RMSSE.

I compare unreconciled base forecasts with reconciled forecasts and report performance both overall and by hierarchy level.

## 7.2 Why level-wise evaluation matters

A reconciliation method can improve aggregate accuracy while worsening bottom-level performance, or vice versa. Therefore, results must be assessed not only overall but also at the total, route, intermediate, and station levels.

```{r echo=FALSE}
############################################################
#  Forecast evaluation
############################################################

# Base forecast accuracy
acc_base <- fc_base |>
  accuracy(test_hier, measures = accuracy_measures)

# Reconciled forecast accuracy
acc_reconciled <- fc_reconciled |>
  accuracy(test_hier, measures = accuracy_measures)

# Summaries by level
acc_base_summary <- summarise_accuracy_by_level(acc_base) |>
  mutate(forecast_family = "Base")

acc_reconciled_summary <- summarise_accuracy_by_level(acc_reconciled) |>
  mutate(forecast_family = "Reconciled")

accuracy_by_level <- bind_rows(acc_base_summary, acc_reconciled_summary)

print(accuracy_by_level)

# Overall summaries across all series
overall_base <- acc_base |>
  group_by(.model) |>
  summarise(
    RMSE  = mean(RMSE, na.rm = TRUE),
    MAE   = mean(MAE, na.rm = TRUE),
    MASE  = mean(MASE, na.rm = TRUE),
    RMSSE = mean(RMSSE, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(forecast_family = "Base")

overall_reconciled <- acc_reconciled |>
  group_by(.model) |>
  summarise(
    RMSE  = mean(RMSE, na.rm = TRUE),
    MAE   = mean(MAE, na.rm = TRUE),
    MASE  = mean(MASE, na.rm = TRUE),
    RMSSE = mean(RMSSE, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(forecast_family = "Reconciled")

overall_accuracy <- bind_rows(overall_base, overall_reconciled) |>
  arrange(RMSSE)

print(overall_accuracy)

# Best reconciled model by hierarchy level
best_reconciled_by_level <- acc_reconciled_summary |>
  group_by(hierarchy_level) |>
  slice_min(order_by = RMSSE, n = 1, with_ties = FALSE) |>
  ungroup()

print(best_reconciled_by_level)
```


# 8. Empirical evaluation results

## 8.1 Base forecast results

Across the unreconciled base forecasts, the [ARIMA / ETS] model achieved the strongest overall holdout performance according to [RMSSE / MASE / RMSE]. The average [chosen metric] was [INSERT VALUE] for [MODEL], compared with [INSERT VALUE] for [OTHER MODEL].

This suggests that [INSERT INTERPRETATION, for example: “the autocorrelation structure captured by ARIMA provided a modest advantage over the smoothing-based specification”].

## 8.2 Reconciled forecast results

Among the reconciled methods, [INSERT BEST METHOD] delivered the strongest overall performance. Its average RMSSE was [INSERT VALUE], compared with [INSERT VALUES] for the other reconciliation methods.

These results indicate that [INSERT INTERPRETATION]. In particular, [INSERT BEST METHOD] appears to provide the best balance between aggregate coherence and predictive accuracy.

## 8.3 Performance by hierarchy level

The level-wise evaluation reveals how the methods trade off local and aggregate accuracy.

### Total level
At the total level, [INSERT METHOD] achieved the best performance, with [INSERT METRIC VALUE]. This indicates that [INSERT INTERPRETATION].

### Route level
At the route level, [INSERT FINDING].

### Route-Direction level
At the route-direction level, [INSERT FINDING].

### Route-Direction-LaneType level
At the route-direction-lane-type level, [INSERT FINDING].

### Station level
At the station level, [INSERT FINDING]. This level is operationally important because station-level accuracy determines the usefulness of the forecasts for local traffic management.

## 8.4 Base versus reconciled forecasts

Comparing unreconciled and reconciled forecasts shows that reconciliation [improved / did not improve] accuracy overall. The main gain came from [INSERT MAIN GAIN], while the main limitation was [INSERT MAIN LIMITATION].

These results show that coherence is not just an accounting constraint. The choice of reconciliation method can materially affect predictive performance.

```{r echo=FALSE}
############################################################
#  Tables and figures for the results section
############################################################

# Plot RMSSE by method and hierarchy level
acc_reconciled_summary |>
  ggplot(aes(x = .model, y = RMSSE)) +
  geom_col() +
  facet_wrap(vars(hierarchy_level), scales = "free_y") +
  coord_flip() +
  labs(
    title = "Reconciled forecast performance by hierarchy level",
    x = "Model",
    y = "Mean RMSSE"
  )

# Top-level forecast plot
fc_total <- fc_reconciled |>
  filter(
    is_aggregated(route),
    is_aggregated(direction),
    is_aggregated(lane_type),
    is_aggregated(station)
  )

actual_total <- bind_rows(train_hier, test_hier) |>
  filter(
    is_aggregated(route),
    is_aggregated(direction),
    is_aggregated(lane_type),
    is_aggregated(station)
  )

autoplot(actual_total, flow) +
  autolayer(fc_total, level = NULL) +
  geom_vline(xintercept = cutoff_time, linetype = "dashed") +
  labs(
    title = "Top-level actuals and reconciled forecasts",
    x = "Time",
    y = "Flow"
  )
```

# 9. Theoretical investigation

This section addresses three theoretical questions:  
(1) Under what conditions is MinT equivalent to bottom-up?  
(2) How does covariance misspecification affect coherence and accuracy?  
(3) When can reconciliation worsen bottom-level accuracy?

## 9.1 When is MinT equivalent to bottom-up?

Let \(\hat{\mathbf y}_h\) denote the vector of base forecasts and let \(\mathbf{S}\) denote the summing matrix. MinT reconciles forecasts according to

\[
\tilde{\mathbf y}_h
=
\mathbf{S}
\left(\mathbf{S}^\top \mathbf{W}_h^{-1}\mathbf{S}\right)^{-1}
\mathbf{S}^\top \mathbf{W}_h^{-1}
\hat{\mathbf y}_h,
\]

where \(\mathbf{W}_h\) is the covariance matrix of base forecast errors.

Bottom-up is a special case in which coherent forecasts are formed solely by aggregating the bottom-level forecasts upward. Therefore, MinT becomes equivalent to bottom-up only in special situations where the optimal weighting scheme places no effective value on non-bottom-level base forecasts. Intuitively, this requires that the aggregate-level forecasts add no information beyond the bottom-level forecasts.

## 9.2 Effect of covariance misspecification

Coherence is imposed algebraically by the reconciliation mapping, so covariance misspecification does not destroy coherence. However, it can reduce accuracy. MinT is optimal only relative to the covariance matrix used in its construction. If that covariance matrix is poorly estimated, the resulting weights may be inefficient or unstable. This is especially relevant in large hierarchies where the number of related series is large relative to the available training history.

This explains why shrinkage-based covariance estimation often performs better in practice than raw sample covariance estimation.

## 9.3 When can reconciliation worsen bottom-level accuracy?

Reconciliation can worsen bottom-level accuracy when the station-level base forecasts are already strong and the reconciliation step pulls them toward noisier or less relevant aggregate information. This can happen when:

1. the bottom-level models already fit the local dynamics well,
2. the covariance matrix used in MinT is poorly estimated,
3. a top-down allocation rule imposes overly smooth aggregate information on heterogeneous bottom-level series.

Therefore, reconciliation should always be evaluated by hierarchy level rather than only at the total level.

## 9.4 Aggregation and variance reduction

Aggregation often reduces noise because idiosyncratic local fluctuations partially cancel out when series are summed. This makes higher-level series easier to forecast in many cases. However, aggregation also discards local structure. Reconciliation is valuable because it attempts to combine the relative stability of upper-level information with the specificity of bottom-level forecasts.

# 10. Sensitivity and robustness analysis

## 10.1 Motivation

Reconciliation performance can depend on both the forecast horizon and the way the hierarchy is defined. To assess the robustness of the main findings, I perform two additional experiments:

1. vary the forecast horizon,
2. change the aggregation structure.

## 10.2 Horizon robustness

I repeat the full forecasting and reconciliation workflow for alternative forecast horizons of 24, 72, and 168 hours. This shows whether the ranking of reconciliation methods changes as the forecasting problem becomes more difficult.

## 10.3 Structural robustness

In the main analysis, I use a strict hierarchy. As an additional robustness check, I also construct an alternative grouped specification by crossing `Route`, `Direction`, and `Lane Type`. This tests whether the main empirical conclusions depend heavily on the exact hierarchy design.

```{r echo=FALSE}
############################################################
#  Sensitivity and robustness analysis
############################################################

# ---------------------------
#  Horizon robustness
# ---------------------------
run_horizon_experiment <- function(panel_data, horizon_hours) {

  cutoff_h <- max(panel_data$timestamp) - hours(horizon_hours)

  split_h <- prepare_hier_split(
    panel_data  = panel_data,
    cutoff_time = cutoff_h
  )

  h_h <- split_h$test_hier |>
    distinct(timestamp) |>
    nrow()

  base_h <- fit_base_suite(split_h$train_hier)
  reconciled_h <- fit_reconciliation_suite(base_h)

  fc_h <- reconciled_h |>
    forecast(h = h_h)

  fc_h |>
    accuracy(split_h$test_hier, measures = accuracy_measures) |>
    add_hierarchy_level() |>
    group_by(.model, hierarchy_level) |>
    summarise(
      RMSE  = mean(RMSE, na.rm = TRUE),
      MAE   = mean(MAE, na.rm = TRUE),
      MASE  = mean(MASE, na.rm = TRUE),
      RMSSE = mean(RMSSE, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(horizon_hours = horizon_hours)
}

robust_h24  <- run_horizon_experiment(traffic_study, 24)
robust_h72  <- run_horizon_experiment(traffic_study, 72)
robust_h168 <- run_horizon_experiment(traffic_study, 168)

robust_horizons <- bind_rows(robust_h24, robust_h72, robust_h168)
print(robust_horizons)

# ---------------------------
# Alternative grouped structure
# ---------------------------
# Uses the SAME leakage-safe main split:
# - train_bottom was imputed from training data only
# - test_bottom contains untouched observed actuals

train_grouped <- train_bottom |>
  aggregate_key(
    (route * direction * lane_type) / station,
    flow = sum(flow)
  )

test_grouped <- test_bottom |>
  aggregate_key(
    (route * direction * lane_type) / station,
    flow = sum(flow)
  )

h_grouped <- test_grouped |>
  distinct(timestamp) |>
  nrow()

base_grouped <- fit_base_suite(train_grouped)
reconciled_grouped <- fit_reconciliation_suite(base_grouped)

fc_grouped <- reconciled_grouped |>
  forecast(h = h_grouped)

grouped_accuracy <- fc_grouped |>
  accuracy(test_grouped, measures = accuracy_measures) |>
  group_by(.model) |>
  summarise(
    RMSE  = mean(RMSE, na.rm = TRUE),
    MAE   = mean(MAE, na.rm = TRUE),
    MASE  = mean(MASE, na.rm = TRUE),
    RMSSE = mean(RMSSE, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(RMSSE)

print(grouped_accuracy)
```


# 11. Robustness results

## 11.1 Horizon sensitivity

The horizon robustness analysis shows that the performance ranking of reconciliation methods [did / did not] remain stable across 24-, 72-, and 168-hour forecast horizons. In particular, [INSERT METHOD] performed best at [INSERT HORIZON], while [INSERT METHOD] became relatively stronger/weaker as the horizon increased.

This suggests that [INSERT INTERPRETATION].

## 11.2 Structural sensitivity

When the strict hierarchy was replaced with an alternative grouped representation, the ranking of methods [remained broadly similar / changed materially]. This indicates that the main conclusions are [structurally robust / sensitive to hierarchy definition].

## 11.3 Implications

Overall, the robustness analysis suggests that [INSERT FINAL ROBUSTNESS TAKEAWAY]. This strengthens the credibility of the preferred reconciliation method for practical use.


# 12. Conclusion

This project developed a coherent hierarchical forecasting system for hourly traffic flow data using the `fpp3` framework in R. The analysis combined station-level series into a multilevel additive hierarchy and compared unreconciled base forecasts with several reconciliation strategies, including bottom-up, top-down, and MinT-based methods.

The study was designed to be both methodologically rigorous and practically relevant. In particular, the workflow avoided data leakage by splitting the sample before any training imputation, imputing only the training data, and reserving the test set for untouched out-of-sample evaluation.

The empirical results show that [INSERT BEST METHOD] provided the best overall balance between coherence and accuracy. More broadly, the analysis demonstrates that reconciliation is not merely a bookkeeping adjustment. The choice of reconciliation method can materially influence forecasting performance, and its effect depends on the hierarchy level, covariance assumptions, and forecast horizon.

Overall, this project shows that hierarchical forecasting is both theoretically rich and operationally important. In a traffic setting, coherent forecasts can improve reporting consistency, network planning, and decision-making across multiple levels of aggregation.
